### 2016-05-23 writing/latex/abstracteng.tex
Henning pres & IS abstract

A permutation is an arrangement of \(n\) objects.
Permutation classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classify mesh patterns of length 2 when avoiding the classical pattern 231.
Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
of sets of the form \(231,m_1\) and \(321,m_2\), as well as discussing possible
future work that could be derived from this work.

>>>>

A permutation is an arrangement of \(n\) objects.
Sets of permutations classified by the avoidance of classical permutation
patterns capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases, are used to completely
classify coincidence amongst families containing a mesh pattern of length \(2\)
and a classical pattern of length \(3\).

Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classify mesh patterns of length \(2\) when avoiding the classical pattern \(231\).
Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
of sets of the form \(231,m_1\) and \(321,m_2\), as well as discussing possible
future work that could be derived from this work.

### 2016-05-23 writing/presentation/coincidences.tex
Henning pres & IS abstract

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they avoid the
    same set of permutations at every length.
  \end{definition}
  \begin{block}{}
    Distinct classical patterns can never be coincident.

    Aim to establish rules that classify coincidences when we
    have one mesh pattern and one classical pattern.
  \end{block}
\end{frame}

@@

\begin{frame}
  \begin{example}
  Considering \(m_1\) again
  \begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        {\only<3-> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        {\only<2> \filldraw (1.66,1.33) circle (3pt);}
        {\only<2> \filldraw (1.33,1.66) circle (3pt);}
    \end{tikzpicture}
  \end{equation*}
  {\only<1-3> \vphantom{This is \(m_2\).}}
  {\only<4-> {}}
  \end{example}
\end{frame}

>>>>

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they are avoided by the
    same set of permutations at every length.
  \end{definition}
  \begin{block}{}
    Distinct classical patterns can never be coincident.

    Coincidences of mesh patterns of length \(2\) are classified.

    Aim to establish rules that classify coincidences when we
    have one mesh pattern and one classical pattern.

    We call the classical pattern a \emph{dominating pattern}.
  \end{block}
\end{frame}

@@

\begin{frame}
  \begin{example}
  Considering \(m_1\) again
  \begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        {\only<3-4> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        {\only<2> \filldraw (1.66,1.33) circle (3pt);}
        {\only<2> \filldraw (1.33,1.66) circle (3pt);}
        {\only<3> \filldraw (2.66,0.33) circle (3pt);}
        {\only<3> \filldraw (2.33,0.66) circle (3pt);}
    \end{tikzpicture}
  \end{equation*}
  {\only<1-3> \vphantom{This is \(m_2\).}}
  {\only<4-> {}}
  \end{example}
\end{frame}

### 2016-05-23 writing/presentation/wilf.tex
Henning pres & IS abstract

 The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
 Catalan numbers.
 \begin{equation}
   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
 \end{equation}
 This gives the sequence
 \begin{equation*}
   1,1,2,5,14,42,132,1430,4862,\dotsc
 \end{equation*}
\end{frame}
\subsection{Set Partitions}
\label{sub:Set Partitions}
\begin{frame}{Set Partitions}
  \begin{block}{Note}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (Claesson \cite{A})
\end{block}
\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}
All permutations in \(\av{\perm{2,3,1}}\) are also in \(\av{q}\).
\end{frame}
\begin{frame}
  Set Partitions can be used to explain certain Wilf-equivalences.
  \begin{example}
    The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
  \begin{equation*}
      m_1 = \pattern{}{1,2}{0/0,0/2,
                      1/0,1/1,
                      2/0,2/1,2/2}
      \text{ and }
      m_2 = \pattern{}{1,2}{0/0,
                      1/0,1/1,1/2,
                      2/0,2/1,2/2}
  \end{equation*}
  Consider containers of the patterns in \(\av{\perm{2,3,1}}\).

@@

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}

>>>>

 The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
 Catalan numbers.
 \begin{equation}
   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
 \end{equation}
 This gives the sequence
 \begin{equation*}
   1,1,2,5,14,42,132,1430,4862,\dotsc
 \end{equation*}
\end{frame}
\subsection{Set Partitions}
\label{sub:Set Partitions}
\begin{frame}{Set Partitions}
  \begin{block}{Note}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (Claesson \cite{A})
\end{block}
\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
\end{example}
All permutations in \(\av{\perm{2,3,1}}\) are also in \(\av{q}\).
\end{frame}
\begin{frame}
  Set Partitions can be used to explain certain Wilf-equivalences.
  \begin{example}
    The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
  \begin{equation*}
      m_1 = \pattern{}{1,2}{0/0,0/2,
                      1/0,1/1,
                      2/0,2/1,2/2}
      \text{ and }
      m_2 = \pattern{}{1,2}{0/0,
                      1/0,1/1,1/2,
                      2/0,2/1,2/2}
  \end{equation*}
  Consider containers of the patterns in \(\av{\perm{2,3,1}}\).

@@

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}
